---
title: "PA1_template"
output: 
  html_document:
    keep_md: true
---

##Loading the data

Reading the file **activity.csv** into a dataframe **dfAct**. This file have to be copied into working directory.

```{r}
dfAct<-read.csv("activity.csv")
```

##Mean total number of steps taken per day

Summarizing number of steps for each day. As a result get the dataframe **StepsByDays** with the variables named **data** and **steps**.

```{r}
StepsByDays<-aggregate(dfAct$steps,list(dfAct$date),sum, na.rm=TRUE)
names(StepsByDays) <- c("date","steps")
```

Ploting the histogram.

```{r}
hist(StepsByDays$steps,xlab="Number of steps", ylab="Days", main = "Total number of steps taken each day", 
       breaks=seq(from=0,to=25000,by=1000), col = "#30D5C8")
```

Counting mean and median values per day.

```{r}
mst <- round(mean(StepsByDays$steps))
mdst <-round(median(StepsByDays$steps))
print(paste("Mean value of steps per day =", mst))
print(paste("Median value of steps per day =", mdst))
```

##The average daily activity pattern

Averaging number of steps for each interval. As a result get the dataframe **StepsByInt** with the variables named **interval** and **steps**.

```{r}
StepsByInt <- aggregate(dfAct$steps,list(dfAct$interval),mean,na.rm=TRUE)
names(StepsByInt) <- c("interval","steps")
```

Ploting the histogram.

```{r}
with(StepsByInt, 
plot(interval,steps, type="l", col="#735184", lwd=2,
     main="Average number of steps on interval"))
```

Counting maximal average number of steps per interval and finding corresponding interval.

```{r}
mxsteps <- max(StepsByInt$steps)
print(paste("Interval witn maximalaverage number of steps =", StepsByInt[StepsByInt$steps==mxsteps,1]))    
print(paste("Maximal average number of steps =", round(mxsteps)))
```

##Imputing missing values

Counting number of the observations with missed value of variable steps.

```{r}
q_na <- length(dfAct$steps[is.na(dfAct$steps)])
print(paste("Total number of missed obsevations =", q_na))
```


Changing NA values in dfAct to mean numbers of steps for this interval. Forming as a result the dataframe **dfActImp** with the variables **steps**, **date**, **interval**.

```{r}
aux <- merge(dfAct, StepsByInt, by = "interval")
aux$steps.x[is.na(aux$steps.x)] <- round(aux$steps.y[is.na(aux$steps.x)])
aux <- aux[order(aux$date,aux$interval),]
dfActImp <- aux[c("steps.x","date","interval")]
names(dfActImp) <- c("steps","date","interval")
```

Get the dataframe **StepsByDaysImp** as analogue of **StepsByDays** but with imputed NA values of steps.

```{r}
StepsByDaysImp<-aggregate(dfActImp$steps,list(dfActImp$date),sum, na.rm=TRUE)
names(StepsByDaysImp) <- c("date","steps")
```

Ploting the histogram for the imputed variant.

```{r}
hist(StepsByDaysImp$steps,xlab="Number of steps", ylab="Days", main = "Total number of steps taken each day (NA imputed)", breaks=seq(from=0,to=25000,by=1000), col = "#9966CC")
```

Counting the new version of mean and median values per day. Imputing missing data impacts on estimates but not so much.

```{r}
mstf <- round(mean(StepsByDaysImp$steps))
mdstf <-round(median(StepsByDaysImp$steps))
print(paste("Mean value of steps per day (NA imputed) =", mstf))
print(paste("Median value of steps per day (NA imputed) =", mdstf))
```

##Differences in activity patterns between weekdays and weekends

Plotting a chart to compare activity in weekends and weekdays.

```{r, results='hide'}
Sys.setlocale(locale = "USA")
```
```{r}
dfActImp$TypeOfDay <- as.factor(ifelse(weekdays((as.Date(dfActImp$date)))%in%c("Saturday","Sunday"),"weekend","weekday"))
library(lattice)
dfActImpWeek <- with(dfActImp, aggregate(steps,list(TypeOfDay,interval),mean))
names(dfActImpWeek) <- c("TypeOfDay","interval", "steps") 
mygraph <- xyplot(steps~interval|TypeOfDay, data = dfActImpWeek, type="l", layout=c(1,2),
                  main = "Effect of weekend on the number of steps",
                  ylab="Number of steps", xlab = "Interval")
plot(mygraph)
```

